### 2023-04-04 main.tex
abs
ii
Update main.tex

\begingroup
% \centering
\begin{center}
 \begingroup\LARGE
\bf Multispecies stirring process
\par\endgroup
 \vspace{3.5em}
 \begingroup\large \bf
% {Rouven Frassea}
 \par\endgroup
\vspace{2em}

@@

\subsection{The infinitesimal generator}

@@

\section{Duality}\label{sectionDuality}
\subsection{Definition}
Consider two Markov processes $(\eta_{t})_{t\geq 0}$ defined on a state space $\Omega$ and $(\xi_{t})_{t\geq 0}$ defined on a state space $\widetilde{\Omega}$. We say that they are dual, with respect to a duality function $D:\Omega\times \widetilde{\Omega}\to \mathbb{R}$, if $\forall \eta\in\Omega$, $\forall \xi\in\widetilde{\Omega}$ and $\forall t> 0$ we have 
    \mathbb{E}_{\eta}\left[D(\eta_{t},\xi)\right]=\mathbb{E}_{\xi}\left[D(\eta,\xi_{t})\right]
where $\mathbb{E}_{\eta}$ denotes the expectation with respect the law of the Markov process $(\eta_{t})_{t\geq 0}$ initialized with the particle configuration $\eta$, whereas $\mathbb{E}_{\xi}$ denotes the expectation with respect to the law of the Markov process $(\xi_{t})_{t\geq 0}$ initialized with the particle configuration $\xi$.
The duality definition can also be formulated as a relation between the generators. Call $\mathcal{L}$ the generator of $(\eta_{t})_{t\geq0}$ and $\widetilde{\mathcal{L}}$ the generator of $(\xi_{t})_{t\geq 0}$, then we say that these two processes are dual with respect to the duality function $D:\Omega\times \widetilde{\Omega}\to \mathbb{R}$ if $\forall \eta\in\Omega$ and $\forall \xi\in\widetilde{\Omega}$
\begin{equation}\label{dualityRelationGenerator}
    \left(\mathcal{L}D(\cdot,\xi)\right)(\eta)=\left(\widetilde{\mathcal{L}}D(\eta,\cdot)\right)(\xi)
In the specific case where $\mathcal{L}=\widetilde{\mathcal{L}}$ we say that the process is self-dual.
\textbf{Reamrk}:
when the state spaces of the dual processes is finite, the generators and the duality function are matrices with elements $\mathcal{L}(\eta,\eta^{'})$, $\widetilde{\mathcal{L}}(\xi,\xi^{'})$ and $D(\eta,\xi)$ for arbitrary $\eta,\eta^{'}\in\Omega$ and $\xi,\xi^{'}\in \widetilde{\Omega}$. Therefore, we can write the duality relation \eqref{dualityRelationGenerator} as 
    \sum_{\eta^{'}\in\,\Omega}\mathcal{L}(\eta,\eta^{'})D(\eta^{'},\xi)=\sum_{\xi^{'}\in\, \widetilde{\Omega}}\widetilde{\mathcal{L}}(\xi,\xi^{'})D(\eta,\xi^{'})
that can be read as
\begin{equation}\label{dualityIntertwines}
    \mathcal{L}D=D\widetilde{\mathcal{L}}^{\,T}
where the superscript $T$ denotes the matrix transposition. Therefore, the duality relation \eqref{dualityIntertwines} is an intertwining between two linear operators $\mathcal{L}$ and $\widetilde{\mathcal{L}}$. Working with the Hamiltonian operators the duality relation \eqref{dualityIntertwines} reads 
\begin{equation}\label{DualityRelation}
    H^{T}D=D\widetilde{H}
\subsection{Duality for the multi-species stirring process}\label{statementDualitySubsection}
In this section we show that the process $(\mathbf{n}(t))_{t\geq 0}$, defined by the generator \eqref{Generator}, is dual to a process $(\bm{\xi}(t))_{t\geq 0}$ taking values in the enlarged state space
\begin{equation}\label{dualStateSpace}
    \widetilde{\Omega}= \bigtimes_{x\in V} \widetilde{\Omega}_{x}\ = \bigtimes_{x\in V} (\Omega_{x}\times \mathbb{N}_{0}^{N-1})
To each site $x\in V$ we associate an ``extra site'', denoted $\widehat{x}$,
where dual particles will accumulate in the course of time. 
We write the configurations $\bm{\xi} \in \widetilde\Omega$  as
    \bm{\xi}=\left(r_{1}^{x},\ldots,r_{N}^{x},m_{2}^{\widehat{x}},\ldots,m_{N}^{\widehat{x}}\right)_{x\in V}
where the component $r_{a}^{x}$ is interpreted as the number of dual particles of type $k\in \{1,\ldots,N\}$ at site $x$, 
and the component $m_{a}^{\widehat{x}}$  gives the number of dual particles of type $k\in \{2,\ldots,N\}$ at 
the extra-site $\widehat{x}$ connected to $x\in V$.
 The  generator of the dual process is 
 \begin{equation}\label{DualGenerator}
    \widetilde{\mathcal{L}}=\sum_{(x,y)\in \mathcal{E}}\omega_{x,y}\mathcal{L}_{x,y}+\sum_{x\in V}\Gamma_{x}\widetilde{\mathcal{L}}_{x}
$\mathcal{L}_{x,y}$ is defined in \eqref{siteGenerator} and, for any function $f:\widetilde{\Omega}\to \mathbb{R}$ 
\begin{equation}\label{siteDualGenerator}
    \widetilde{\mathcal{L}}_{x}f(\bm{\xi})=\sum_{i=1}^{N}\alpha_{i}^{x}\sum_{a=2}^{N}r_{a}^{x}\left(f(\bm{\xi}-\delta_{a}^{x}+\delta_{1}^{x}+\delta_{a}^{\widehat{x}})-f(\bm{\xi})\right)
and the duality function is 
\begin{equation}\label{dualityElements}
	D(\bm{n},\bm{\xi})=\prod_{x\in V}\left(\frac{(2j-\sum_{a=2}^{N}r_{a}^{x})!}{\nu!}\prod_{a=2}^{N}\frac{n_{a}^{x}!}{(n_{a}^{x}-r_{a}^{x})!}\left(\rho_{a}^{x}\right)^{m_{a}^{\widehat{x}}}\,\right)
where we denote the \textit{density} of the species $k\in \{2,\ldots,N\}$ at the reservoir connected with the site $x\in V$  by 
\rho_{a}^{x}:=\frac{\alpha_{a}^{x}}{\sum_{i=1}^{N}\alpha_{i}^{x}}
On one hand, the edge part of the dual generator \eqref{DualGenerator} is a copy of \eqref{edgeGenerator}, therefore it performs the stirring dynamics on the graph. On the other hand, the dual generator \eqref{siteDualGenerator} replaces a particle of any type $k\in\{2,\ldots,N\}$ at site $x$ with a particle of type $1$ and creates a particle of the same type $k$ at the extra-site $\widehat{x}$. This last transition is performed with rate $\sum_{i=1}^{N}\alpha_{i}^{x}r_{a}^{x}$. This means that eventually the dual process voids the graph, putting all the dual particles of species $\{2,\ldots,N\}$ in the extra-sites. In other words the extra-sites play the role of absorbing boundaries. 
\newline \newline
\textbf{Reamrk}: in the reversible situation, i.e. when $\forall x\in V$ $\alpha_{a}^{x}=\alpha_{a}$, the expectation  of the duality function  $D(\bm{n},\bm{\xi})$ with   $\bm{n}$ distributed as  $\mu_{rev} = \bigtimes_{x\in V}\text{Multinomial}\left(\twoj, p_{1},\ldots,p_{N}\right)$ is
\mathbb{E}_{\mu^{rev}}\left[D(\bm{n},\bm{\xi})\right]=\prod_{a=2}^{N}\left(\rho_{a}^{x}\right)^{\sum_{x\in V}r_{a}^{x}+\sum_{x\in V}m_{a}^{\widehat{x}}}\qquad \forall \bm{n}\in \Omega,\quad\forall \bm{\xi}\in \widetilde{\Omega}

@@

The configuration space  \eqref{dualStateSpace} of the dual process is the collection of vectors

@@

% However, they would lead to duality matrices where the number of dual particles is greater than the number of original ones, loosing the simplification of the dynamics of the dual process. 
 \begin{flushright}
     $\square$
 \end{flushright}
 \paragraph{Proof of \eqref{siteDualRelation}.} To prove \eqref{siteDualRelation} we need to transform via the Hadamard formula the transposed of the site Hamiltonian \eqref{siteDualHamiltonian} and then to  introduce properly a creation operator acting on an extra-site $\widehat{x}$. % that we connect to every site $x\in V$ of the graph $G$. \\
 Considering $A,B\in gl(N)$, the Hadamard formula reads 
 \begin{equation}\label{HadamardFormula}
     \exp{(-B)}A\exp{(B)}=A-\left[B,A\right]+\frac{1}{2!}\left[B,\left[B,A\right]\right]-\frac{1}{3!}\left[B,\left[B,\left[B,A\right]\right]\right]+\ldots
 \end{equation}
In the followin we need the application of the Hadamard formula with $B=E$ defined in \eqref{equationE} and with $A$ equal to some of the generator $E_{ab}$ of the Lie algebra. Therefore we compute 
\begin{enumerate}
    \item for $A=E_{b 1}$ with $b\in \{2,\ldots,N\}$, we obtain 
    \begin{equation}\label{HT_El1}
        \exp{(-E)}E_{b 1}\exp{(E)}=E_{b 1}
    \end{equation}
    because $E_{b 1}$ commutes with $E$
    \item for $A=E_{bb}$ with $b \in \{2,\ldots,N\}$, we obtain 
    \begin{equation}\label{HT_Ell}
        \exp{(-E)}E_{b b}\exp{(E)}=E_{b b}+E_{b 1}
    \end{equation}
Indeed, using \eqref{eq:comgl} we have 
    \begin{equation}
       \left[E, E_{bb}\right]=\sum_{a=2}^{N}\left(E_{ab}\delta_{b 1}-E_{b 1}\delta_{ab}\right)=-E_{l1}
    \end{equation}
    Inserting the above commutator in \eqref{HadamardFormula}, we obtain \eqref{HT_Ell}.
    \item for $A=E_{1b}$ with $b\in \{2,\ldots,N\}$, we obtain 
    \begin{equation}\label{HT-E1l}
        \exp{(-E)}E_{1 b}\exp{(E)}=E_{1b}+E_{11}-\sum_{j=2}^{N}\left(E_{j1}+E_{jb}\right)
    \end{equation}
  \com{Why indices have names $b, k, j, a, b$? Fix one set for the generators}

    Indeed, using \eqref{eq:comgl} we have 
    \begin{equation}
[E,E_{1b}]=\sum_{a=2}^{N}\left(E_{ab}\delta_{11}-E_{11}\delta_{ab}\right)=\sum_{j=2}^{N}E_{jb}-E_{11};
\end{equation}
and 
\begin{equation}
\begin{split}
\left[E,[E,E_{1b}]\right]&=\sum_{b=2}^{N}\sum_{j=2}^{N}\left[E_{b1},E_{jb}\right]-\sum_{b=2}^{N}\left[E_{b1},E_{11}\right]
\\&=
\sum_{j,b=2}^{N}\left(E_{bb}\delta_{j1}-E_{j1}\delta_{bb}\right)-\sum_{b=2}^{N}\left(E_{b1}\delta_{11}-E_{11}\delta_{b1}\right)
\\=&
-2\sum_{j=2}^{N}E_{j1};
\end{split}
\end{equation}
Inserting the above commutators in \eqref{HadamardFormula}, we obtain \eqref{HT-E1l}.
\item for $A=E_{ab}$ with $k,b \in \{2,\ldots,N\}$ we obtain 
\begin{equation}\label{HT-Ekl}
    \exp{(-E)}E_{b a}\exp{(E)}=E_{ab}+E_{b 1}
\end{equation}
   Indeed, using \eqref{eq:comgl} we have 
\begin{equation}
[E,E_{b a}]=\sum_{a=2}^{N}\left(E_{aa}\delta_{b 1}-E_{b 1}\delta_{aa}\right)=-E_{b 1};
\end{equation}
\begin{equation}
[E,E_{ab}]=\sum_{a=2}^{N}\left(E_{ab}\delta_{a1}-E_{a1}\delta_{b a}\right)=-E_{a1};
\end{equation}
Inserting the above commutator in \eqref{HadamardFormula}, we obtain \eqref{HT-Ekl}.
\item for $A=E_{11}$ we obtain 
\begin{equation}\label{HT-E11}
    \exp{(-E)}E_{11}\exp{(E)}=E_{11}-\sum_{j=2}^{N}E_{j1}
\end{equation}
  Indeed, using \eqref{eq:comgl} we have
[E,E_{11}]=\sum_{a=2}^{N}\left(E_{a1}\delta_{11}-E_{1 1}\delta_{a1}\right)=\sum_{j=2}^{N}E_{j1};

@@

H_{x}^{T}=\sum_{a,l=0}^{N}\alpha_{a}^{x}\left(E_{ab}^{x}-E_{bb}^{x}\right)^{T}=R_{x}\sum_{a,b=0}^{N}\alpha_{a}^{x}\left(E_{b a}^{x}-E_{bb}^{x}\right)R_{x}^{-1}

@@

In this section we specialize the multi-species stirring process to the geometry of the one dimensional chain with sites $\{1,\ldots,L\}$ where two reservoirs are connected with the end site $1$ and $L$. These boundaries, exchanges particle with the external environment, putting the chain out of equilibrium. To use integrability is better to work with Hamiltonians. The Hamiltonian that we consider here is obtained by \eqref{OriginalHamiltonian} assuming that the conducatances are

@@

We will denote the element of the representation of the Lie algebra by lowercase letters, i.e. $e_{ab}$, since we are in the lowest spin case (the fundamental representation of algebra). \\
The state space of the chain is 
    \Omega=\bigtimes_{x=1}^{L}\Omega_{x}
	\Omega_{x}=\left\{|n\rangle=|n_{1},\ldots,n_{N}\rangle\;:\;n\in\mathbb{N}_0^N\;\;|n|=1\right\}

@@

Moreover, for a fixed site $x\in \{1,\ldots,L\}$ we have 

@@

	P=\sum_{a,b=1}^Ne_{ab}\otimes e_{ba}

@@

	\sum_{a=2}^{N}\alpha_{a}=1,\qquad\sum_{a=2}^{N}\beta_{a}=1

@@

	\lambda_{a}=\alpha_{a}-\beta_{a}\quad\text{with}\quad \sum_{a=2}^{N}\lambda_{a}=0\,.
Under these assumption the Hamiltonian \eqref{hamiltonian} is integrable and a matrix product anstz (MPA) for the non-equilibrium steady state is available (see\cite{vanicat2017exact}). We recall the main points of this ansatz. \\
 Calling $|\mathbb{P}(t)\rangle$ a vector that contains the probability of every configuration of the chain with Hamiltonian \eqref{hamiltonian} at time $t\geq 0$, its evolution equation is given by 
    \frac{d|\mathbb{P}(t)}{dt}=H|\mathbb{P}(t)\rangle
We define the steady state of $H$ defined in \eqref{hamiltonian} a vector, denoted by $|\Psi\rangle$ that contains the probabilities of every configuration of the chain with respect to the stationary distribution, i.e. the probability distribution that the chain reaches when $t\to \infty$. We observe that, since the Markov chain with Hamiltonian \eqref{hamiltonian} is positive recurrent there exists a unique stationary distribution and, moreover, this is distribution is reached when $t\to \infty$ starting for every arbitrary initial distribution. 

@@

where $X_{a}$ for $a=1,\ldots,N$ are operators on an auxiliary space $V_{0}$ while $|V\rangle\in V_{0}$ and $\langle W|\in V_{0}^{*}$ ( where $V^{*}$ denotes the algebraic dual space), such that $\langle W|V\rangle=1$. $\langle W|$ and $|V\rangle$ are sometimes called boundary vectors. These operators $(X_{a})_{a\in \{1,\ldots,N\}}$ fulfil the commutators

@@

\textbf{Reamrk}: only $N-1$ of \eqref{leftBoundary} and \eqref{rightBoundary} are independent, indeed because of \eqref{ratesConditions} and \eqref{lambdaConditions}

@@

	gives $0=0$. Similar computations can be done for the right boundary.
\subsection{The exact steady state for the multi-species stirring process on a chain}\label{subsection-exact}
We aim to find an explicit for for the steady state $|\Psi\rangle$. The MPA states the form of the steady state in function of an abstract algebra of operators. If an explicit representation of $(X_{a})_{a=1,\ldots,N}$ was available, it would be possible to compute explicitly all the correlations and currents of particles for the Hamiltonian \eqref{hamiltonian}. Another way of computing this steady state, without passing through a representation, is using the commutators \eqref{bula} and the actions \eqref{leftBoundary} and \eqref{rightBoundary}. For the one and two point correlation and for the currents this has been done in \cite{vanicat2017exact}. Inspired by this last reference, in this subsection, we aim to find exact formulas for the steady state of \eqref{hamiltonian} using only the \eqref{bula}, \eqref{leftBoundary} and \eqref{rightBoundary}. In subsection \eqref{correlation-section}, this result will leads to exact correlation for arbitrary number of particles of type $\{2,\ldots,N\}$. This aim will be achieved by the help of a similarity transformation $\mathcal{S}$ that makes commutation relation simpler, and allows to find the steady state of a transformed Hamiltonian, that will be denoted by $\mathbb{H}$, with lower-triangular left boundary and diagonal right boundary. Therefore, inverting this similarity, one can retrieve the steady state of \eqref{hamiltonian}. A fundamental 'ingredient' of these computation is duality, that has been proved in section~\ref{sectionDuality}. Indeed, the similarity transformation $\mathcal{S}$ relies on an "intermediate transformation" $V$ (see \eqref{transformationV}) that is, up to the diagonal operator $R$, the duality matrix. \com{clarify}

@@

\sum_{i=1\,:\, i\neq q_{1},\ldots,q_{b}}^{L}\sum_{c_{i}=0}^{1-\delta_{s_{x},1}}\sum_{p=1}^{b}\sum_{c_{q_{p}}=0}^{1}\frac{\Gamma(2+L-\sum_{a=1}^{L}c_{a})}{\Gamma(2+L)}
  \prod_{i=1\,:\, i\neq q_{1},\ldots,q_{b}}^{L}\left(\lambda_{s_{x}}\left(2+L-x-\sum_{j=x}^{L}c_{j}\right)\right)^{c_{i}}

@@

 \beta_{s_{x}}^{(1-c_{i})(1-\delta_{s_{x},1})}

>>>>

\begingroup
% \centering
\begin{center}
 \begingroup\LARGE
\bf Multispecies stirring process
\par\endgroup
 \vspace{3.5em}
 \begingroup\large \bf
% {Rouven Frassek}
 \par\endgroup
\vspace{2em}

@@

\subsection{The process generator}

@@

\section{Duality}\label{sectionDuality}
\subsection{Definition}
Consider two Markov processes $(\eta_{t})_{t\geq 0}$ defined on a state space $\Omega$ and $(\xi_{t})_{t\geq 0}$ defined on a state space $\widetilde{\Omega}$. We say that they are dual, with respect to a duality function $D:\Omega\times \widetilde{\Omega}\to \mathbb{R}$, if $\forall \eta\in\Omega$, $\forall \xi\in\widetilde{\Omega}$ and $\forall t> 0$ we have 
\begin{equation}
    \mathbb{E}_{\eta}\left[D(\eta_{t},\xi)\right]=\mathbb{E}_{\xi}\left[D(\eta,\xi_{t})\right]
\end{equation}
where $\mathbb{E}_{\eta}$ denotes the expectation with respect the law of the Markov process $(\eta_{t})_{t\geq 0}$ initialized with the particle configuration $\eta$, whereas $\mathbb{E}_{\xi}$ denotes the expectation with respect to the law of the Markov process $(\xi_{t})_{t\geq 0}$ initialized with the particle configuration $\xi$.
The duality definition can also be formulated as a relation between the generators. Call $\mathcal{L}$ the generator of $(\eta_{t})_{t\geq0}$ and $\widetilde{\mathcal{L}}$ the generator of $(\xi_{t})_{t\geq 0}$, then we say that these two processes are dual with respect to the duality function $D:\Omega\times \widetilde{\Omega}\to \mathbb{R}$ if $\forall \eta\in\Omega$ and $\forall \xi\in\widetilde{\Omega}$
\begin{equation}\label{dualityRelationGenerator}
    \left(\mathcal{L}D(\cdot,\xi)\right)(\eta)=\left(\widetilde{\mathcal{L}}D(\eta,\cdot)\right)(\xi)
\end{equation}
In the specific case where $\mathcal{L}=\widetilde{\mathcal{L}}$ we say that the process is self-dual.
\newline
\newline
\textbf{Reamrk}:
when the state spaces of the dual processes is finite, the generators and the duality function are matrices with elements $\mathcal{L}(\eta,\eta^{'})$, $\widetilde{\mathcal{L}}(\xi,\xi^{'})$ and $D(\eta,\xi)$ for arbitrary $\eta,\eta^{'}\in\Omega$ and $\xi,\xi^{'}\in \widetilde{\Omega}$. Therefore, we can write the duality relation \eqref{dualityRelationGenerator} as 
\begin{equation}
    \sum_{\eta^{'}\in\,\Omega}\mathcal{L}(\eta,\eta^{'})D(\eta^{'},\xi)=\sum_{\xi^{'}\in\, \widetilde{\Omega}}\widetilde{\mathcal{L}}(\xi,\xi^{'})D(\eta,\xi^{'})
\end{equation}
that can be read as
\begin{equation}\label{dualityIntertwines}
    \mathcal{L}D=D\widetilde{\mathcal{L}}^{\,T}
\end{equation}
where the superscript $T$ denotes the matrix transposition. Therefore, the duality relation \eqref{dualityIntertwines} is an intertwining between two linear operators $\mathcal{L}$ and $\widetilde{\mathcal{L}}$. Working with the Hamiltonian operators the duality relation \eqref{dualityIntertwines} reads 
\begin{equation}\label{DualityRelation}
    H^{T}D=D\widetilde{H}
\end{equation}
\subsection{Duality for the multi-species stirring process}\label{statementDualitySubsection}
In this section we formulate duality for the multi-species stirring process $(\mathbf{n}(t))_{t\geq 0}$, defined by the generator \eqref{Generator}.
The dual  process $(\bm{\xi}(t))_{t\geq 0}$ takes values in the enlarged state space
\begin{equation}\label{dualStateSpace}
    \widetilde{\Omega}= \bigtimes_{x\in V} \widetilde{\Omega}_{x}\ = \bigtimes_{x\in V} (\Omega_{x}\times \mathbb{N}_{0}^{N-1})
\end{equation}
To each site $x\in V$ we associate an ``extra site'', denoted $\widehat{x}$,
where dual particles will accumulate in the course of time. 
We write the configurations $\bm{\xi} \in \widetilde\Omega$  as
\begin{equation}
    \bm{\xi}=\left(r_{1}^{x},\ldots,r_{N}^{x},m_{2}^{\widehat{x}},\ldots,m_{N}^{\widehat{x}}\right)_{x\in V}
\end{equation}
where the component $r_{a}^{x}$ is interpreted as the number of dual particles of type $a\in \{1,\ldots,N\}$ at site $x$, 
and the component $m_{a}^{\widehat{x}}$  gives the number of dual particles of type $a\in \{2,\ldots,N\}$ at 
the extra-site $\widehat{x}$ connected to $x\in V$.

The duality function is given by
\begin{equation}\label{dualityElements}
	D(\bm{n},\bm{\xi})=\prod_{x\in V}\left(\frac{(\nu -\sum_{a=2}^{N}r_{a}^{x})!}{\nu!}\prod_{a=2}^{N}\frac{n_{a}^{x}!}{(n_{a}^{x}-r_{a}^{x})!}\left(\rho_{a}^{x}\right)^{m_{a}^{\widehat{x}}}\,\right)
\end{equation}
where we denote the \textit{density} of the species $a\in \{2,\ldots,N\}$ imposed by the reservoir at site $x\in V$  by 
\begin{equation}
\rho_{a}^{x}:=\frac{\alpha_{a}^{x}}{\sum_{i=1}^{N}\alpha_{i}^{x}}
\end{equation}
The dual process is defined by its generator, which reads
 \begin{equation}\label{DualGenerator}
    \widetilde{\mathcal{L}}=\sum_{(x,y)\in \mathcal{E}}\omega_{x,y}\mathcal{L}_{x,y}+\sum_{x\in V}\Gamma_{x}\widetilde{\mathcal{L}}_{x}
\end{equation}
where 
$\mathcal{L}_{x,y}$ is defined in \eqref{edgeGenerator} and, for any function $f:\widetilde{\Omega}\to \mathbb{R}$ 
\begin{equation}\label{siteDualGenerator}
    \widetilde{\mathcal{L}}_{x}f(\bm{\xi})=\sum_{i=1}^{N}\alpha_{i}^{x}\sum_{a=2}^{N}r_{a}^{x}\left(f(\bm{\xi}-\delta_{a}^{x}+\delta_{1}^{x}+\delta_{a}^{\widehat{x}})-f(\bm{\xi})\right)
\end{equation}
\newline
Therefore the dual dynamics is described as follows. On one hand, the edge part $\mathcal{L}_{x,y}$ of the dual generator gives rise to  the multi-species stirring dynamics on the graph. On the other hand, the site
part $\widetilde{\mathcal{L}}_{x}$ of the dual generator replaces a particle of any type $a\in\{2,\ldots,N\}$ at site $x$ with a particle of type $1$ and creates a particle of the same type $a$ at the extra-site $\widehat{x}$. This last transition is performed with rate $\sum_{i=1}^{N}\alpha_{i}^{x}r_{a}^{x}$. This means that eventually the dual process voids the graph, putting all the dual particles of species $\{2,\ldots,N\}$ in the extra-sites. In other words the extra-sites play the role of absorbing boundaries. 
\newline \newline
\textbf{Remark}: in the reversible situation, i.e. when $\forall x\in V$ $\rho_{a}^{x}=\rho_{a}$, the expectation  of the duality function  $D(\bm{n},\bm{\xi})$ with   $\bm{n}$ distributed as  $\mu_{rev} = \bigtimes_{x\in V}\text{Multinomial}\left(\twoj, \rho_{1},\ldots,\rho_{N}\right)$ is
\begin{equation}
\mathbb{E}_{\mu^{rev}}\left[D(\bm{n},\bm{\xi})\right]=\prod_{a=2}^{N}\left(\rho_{a}\right)^{\sum_{x\in V}r_{a}^{x}+\sum_{x\in V}m_{a}^{\widehat{x}}}\qquad \forall \bm{n}\in \Omega,\quad\forall \bm{\xi}\in \widetilde{\Omega}
\end{equation}

@@

To each configuration $\bm{\xi}$ of the configuration space  \eqref{dualStateSpace} of the dual process we associate the vector

@@

% However, they would lead to duality matrices where the number of dual particles is greater than the number of original ones, loosing the simplification of the dynamics of the dual process. 
 \begin{flushright}
     $\square$
 \end{flushright}
 \paragraph{Proof of \eqref{siteDualRelation}.} To prove \eqref{siteDualRelation} we need to transform via the Hadamard formula the transposed of the site Hamiltonian \eqref{siteDualHamiltonian} and then to  introduce properly a creation operator acting on an extra-site $\widehat{x}$. % that we connect to every site $x\in V$ of the graph $G$. \\
 Considering $A,B\in gl(N)$, the Hadamard formula reads 
 \begin{equation}\label{HadamardFormula}
     \exp{(-B)}A\exp{(B)}=A-\left[B,A\right]+\frac{1}{2!}\left[B,\left[B,A\right]\right]-\frac{1}{3!}\left[B,\left[B,\left[B,A\right]\right]\right]+\ldots
 \end{equation}
In the followin we need the application of the Hadamard formula with $B=E$ defined in \eqref{equationE} and with $A$ equal to some of the generator $E_{ab}$ of the Lie algebra. Therefore we compute 
\begin{enumerate}
    \item for $A=E_{b 1}$ with $b\in \{2,\ldots,N\}$, we obtain 
    \begin{equation}\label{HT_El1}
        \exp{(-E)}E_{b 1}\exp{(E)}=E_{b 1}
    \end{equation}
    because $E_{b 1}$ commutes with $E$
    \item for $A=E_{bb}$ with $b \in \{2,\ldots,N\}$, we obtain 
    \begin{equation}\label{HT_Ell}
        \exp{(-E)}E_{b b}\exp{(E)}=E_{b b}+E_{b 1}
    \end{equation}
Indeed, using \eqref{eq:comgl} we have 
    \begin{equation}
       \left[E, E_{bb}\right]=\sum_{a=2}^{N}\left(E_{ab}\delta_{b 1}-E_{b 1}\delta_{ab}\right)=-E_{b1}
    \end{equation}
    Inserting the above commutator in \eqref{HadamardFormula}, we obtain \eqref{HT_Ell}.
    \item for $A=E_{1b}$ with $b\in \{2,\ldots,N\}$, we obtain 
    \begin{equation}\label{HT-E1l}
        \exp{(-E)}E_{1 b}\exp{(E)}=E_{1b}+E_{11}-\sum_{a=2}^{N}\left(E_{a1}+E_{ab}\right)
    \end{equation}
  %\com{Why indices have names $b, k, j, a, b$? Fix one set for the generators}

    Indeed, using \eqref{eq:comgl} we have 
    \begin{equation}
[E,E_{1b}]=\sum_{a=2}^{N}\left(E_{ab}\delta_{11}-E_{11}\delta_{ab}\right)= (\sum_{a=2}^{N}E_{ab} ) -E_{11};
\end{equation}
and 
\begin{equation}
\begin{split}
\left[E,[E,E_{1b}]\right]&=\sum_{b=2}^{N}\sum_{a=2}^{N}\left[E_{b1},E_{ab}\right]-\sum_{b=2}^{N}\left[E_{b1},E_{11}\right]
\\&=
\sum_{a,b=2}^{N}\left(E_{bb}\delta_{a1}-E_{a1}\delta_{bb}\right)-\sum_{b=2}^{N}\left(E_{b1}\delta_{11}-E_{11}\delta_{b1}\right)
\\=&
-2\sum_{a=2}^{N}E_{a1};
\end{split}
\end{equation}
Inserting the above commutators in \eqref{HadamardFormula}, we obtain \eqref{HT-E1l}.
\item for $A=E_{ab}$ with $a,b \in \{2,\ldots,N\}$ we obtain 
\begin{equation}\label{HT-Ekl}
    \exp{(-E)}E_{a b}\exp{(E)}=E_{ab}+E_{a 1}
\end{equation}
   Indeed, using \eqref{eq:comgl} we have 
\begin{equation}
[E,E_{ab}]=\sum_{c=2}^{N}\left(E_{cb}\delta_{a1}-E_{a1}\delta_{c b}\right)=-E_{a1};
\end{equation}
Inserting the above commutator in \eqref{HadamardFormula}, we obtain \eqref{HT-Ekl}.
\item for $A=E_{11}$ we obtain 
\begin{equation}\label{HT-E11}
    \exp{(-E)}E_{11}\exp{(E)}=E_{11}-\sum_{a=2}^{N}E_{a1}
\end{equation}
  Indeed, using \eqref{eq:comgl} we have
[E,E_{11}]=\sum_{a=2}^{N}\left(E_{a1}\delta_{11}-E_{1 1}\delta_{a1}\right)=\sum_{a=2}^{N}E_{a1};

@@

H_{x}^{T}=\sum_{a,b=0}^{N}\alpha_{a}^{x}\left(E_{ab}^{x}-E_{bb}^{x}\right)^{T}=R_{x}\sum_{a,b=0}^{N}\alpha_{a}^{x}\left(E_{b a}^{x}-E_{bb}^{x}\right)R_{x}^{-1}

@@

In this section we specialize the multi-species stirring process to the geometry of the one dimensional chain with sites $\{1,\ldots,L\}$ where two reservoirs act at the boundary sites $1$ and $L$. These boundaries, exchanging particle with the external environment, put the chain out of equilibrium. The Hamiltonian that we consider here is obtained by \eqref{OriginalHamiltonian} assuming that the conductances are

@@

{\color{red} We will denote the element of the representation of the Lie algebra by lowercase letters, i.e. $e_{ab}$, since we are in the lowest spin case (the fundamental representation of algebra).} \\

@@

Here $\mathcal{H}_{x,x+1}$  is a copy (acting on the edge $x,x+1$) of the two-site Hamiltonian

@@

	P=\sum_{a,b=1}^Ne_{ab}\otimes e_{ba}.
The action of $\mathcal{H}$ amounts to the permutation of the two sites, i.e.

@@

	\sum_{a=1}^{N}\alpha_{a}=1,\qquad\sum_{a=1}^{N}\beta_{a}=1

@@

	\lambda_{a}=\alpha_{a}-\beta_{a}\quad\text{with}\quad \sum_{a=1}^{N}\lambda_{a}=0\,.
Under these assumption the Hamiltonian \eqref{hamiltonian} is integrable and a matrix product ansatz (MPA) for the non-equilibrium steady state is available (see\cite{vanicat2017exact}). 
 Calling $|{P}(t)\rangle$ a vector that contains the probability of every configuration of the chain with Hamiltonian \eqref{hamiltonian} at time $t\geq 0$, its evolution equation is given by 
 the master equation
    \frac{d|{P}(t)\rangle}{dt}=H|{P}(t)\rangle
We denote by $|\Psi\rangle$ that contains the probabilities of every configuration of the chain with respect to the stationary distribution, i.e. the probability distribution that the chain reaches when $t\to \infty$. We observe that, since the Markov chain with Hamiltonian \eqref{hamiltonian} is positive recurrent there exists a unique stationary distribution and, moreover, this is distribution is reached when $t\to \infty$ starting for every arbitrary initial distribution. 

@@

where the operators $(X_{a})_{a\in \{1,\ldots,N\}}$ fulfill the commutators

@@

\textbf{Remark}: only $N-1$ of the $N$ equations \eqref{leftBoundary}  are independent. This can be seen by summing them 

@@

	and using  \eqref{ratesConditions} and \eqref{lambdaConditions}.
 Similarly for the right boundary  \eqref{rightBoundary} there are only  $N-1$ independent equations.
\subsection{The explicit form of the steady state}\label{subsection-exact}
We aim to find an explicit for for the steady state $|\Psi\rangle$. The MPA gives the steady state in terms of an abstract algebra of operators. If an explicit representation of $(X_{a})_{a=1,\ldots,N}$ was available, it would be possible to compute explicitly all the correlations and currents of particles for the Hamiltonian \eqref{hamiltonian}. Another way of computing this steady state, without passing through a representation, is using the commutators \eqref{bula} and the actions \eqref{leftBoundary} and \eqref{rightBoundary}. For the one and two point correlation and for the currents this has been done in \cite{vanicat2017exact}. Inspired by this last reference, in this subsection, we aim to find exact formulas for the steady state  $|\Psi\rangle$ using only the \eqref{bula}, \eqref{leftBoundary} and \eqref{rightBoundary}. In subsection \ref{correlation-section}, this result will leads to exact correlation for arbitrary number of particles of type $\{2,\ldots,N\}$. This aim will be achieved by the help of a similarity transformation $\mathcal{S}$ that makes commutation relations simpler, and allows to find the steady state $|\widetilde\Psi\rangle$ of a transformed Hamiltonian, denoted  $\mathbb{H}$, with lower-triangular left boundary and diagonal right boundary. Therefore, inverting this similarity, one can retrieve the steady state of \eqref{hamiltonian}. A fundamental 'ingredient' of these computation is duality, that has been proved in section~\ref{sectionDuality}. Indeed, the similarity transformation $\mathcal{S}$ relies on an "intermediate transformation" $V$ (see \eqref{transformationV}) that is, up to the diagonal operator $R$, the duality matrix. \com{clarify}

@@

\sum_{x=1\,:\, x\neq q_{1},\ldots,q_{b}}^{L}\sum_{c_{x}=0}^{1-\delta_{s_{x},1}}\sum_{p=1}^{b}\sum_{c_{q_{p}}=0}^{1}\frac{\Gamma(2+L-\sum_{a=1}^{L}c_{a})}{\Gamma(2+L)}
  \prod_{x=1\,:\, x\neq q_{1},\ldots,q_{b}}^{L}\left(\lambda_{s_{x}}\left(2+L-x-\sum_{j=x}^{L}c_{j}\right)\right)^{c_{x}}

@@

 \beta_{s_{x}}^{(1-c_{x})(1-\delta_{s_{x},1})}
